---
title: "Mapping Chinese Churches in Japanese Occupied Hong Kong"
date: "2026-08-13"
author: "Clarissa Lo"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## 1. Introduction & Packages Loading

This report presents the spatial distribution and social roles of Chinese churches in Hong Kong during the Japanese Occupation using 'leaflet'.

```{r load-packages}
library(leaflet)
library(tidyverse)
library(sf)
library(ggplot2)
library(dplyr)

hk_districts <- st_read("hksar_18_district_boundary.json", quiet = TRUE)
```

## 2. Data Preparation & Coordinates Processing

Read 'wartime_churches.csv' and construct English HTML popups for the map interactivity.

```{r process-data}
df <- read_csv("wartime_churches.csv")

df_map <- df %>%


  mutate(
    new_longitude = as.numeric(new_longitude),
    origin_longitude = as.numeric(origin_longitude),
    new_latitude = as.numeric(new_latitude),
    origin_latitude = as.numeric(origin_latitude)
  ) %>%
 
  mutate(
    lng = if_else(!is.na(new_longitude), new_longitude, origin_longitude),
    lat = if_else(!is.na(new_latitude), new_latitude, origin_latitude)
  ) %>%
  
  filter(!is.na(lng) & !is.na(lat)) %>%
 
  mutate(
    has_safety_net = ifelse(
      !is.na(safety_net_service) & safety_net_service != "N/A" & safety_net_service != "None", 
      "Provided", 
      "None / No Record"
    ),
    marker_size = ifelse(has_safety_net == "Provided", 10, 5),
    popup_content = paste0(
  "<b>", church_name_en, "</b> (", church_name_ch, ")<br>",
  "<b>Address:</b> ", address_cn, "<br>",
  "<b>Wartime Status:</b> ", ifelse(is.na(wartime_status), "No Record", wartime_status), "<br>",
  "<b>Safety Net Service:</b> ", ifelse(is.na(safety_net_service) | safety_net_service == "N/A", "None / No Record", safety_net_service), "<br>",
  "<b>Japanese Policy Impact:</b> ", ifelse(is.na(japanese_policy_impact), "No Record", japanese_policy_impact), "<br><br>",
  "<b>Description (EN):</b> ", ifelse(is.na(description_en) | description_en == "", "N/A", description_en), "<br>",
  "<b>Description (ZH):</b> ", ifelse(is.na(description_ch) | description_ch == "", "無紀錄", description_ch)
)
  )
```

## 3. Interactive Map of Wartime Churches

The markers are color-coded based on each church's **Wartime Status ('wartime_status')**:

```{r render-map}
# Define color palette for wartime status 
pal_status <- colorFactor(palette = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd"),
  domain = df_map$wartime_status)

leaflet(df_map) %>% 
  addTiles() %>% 
  setView(lng = 114.168, lat = 22.319, zoom = 12) %>%
addPolygons(
    data = hk_districts,
    fillColor = "transparent", 
    color = "#444444",          
    weight = 1,                 
    dashArray = "3",
    opacity = 0.7
  ) %>% addCircleMarkers(
    lng = ~lng, 
    lat = ~lat, 
    radius = ~marker_size,                  
    color = ~pal_status(wartime_status),    
    stroke = TRUE,                          
    weight = 1,
    fillOpacity = 0.8, 
    popup = ~popup_content, 
    label = ~church_name_en
  ) %>% 
  # Sample 1： Wartime Status (color)
  addLegend(
    position = "bottomright", 
    pal = pal_status, 
    values = ~wartime_status, 
    title = "Wartime Status (Color)", 
    opacity = 1
  ) %>%
  # Sample 2： Safety Net Service (size difference)
  addLegend(
    position = "bottomleft",
    colors = c("gray", "gray"),
    labels = c("Large Circle: Safety Net Provided", "Small Circle: No Safety Net / No Record"),
    title = "Safety Net Service (Size)",
    opacity = 1)
```

```{r}
pie_data <- df %>%
  mutate(safety_net_service = ifelse(is.na(safety_net_service) | safety_net_service == "", "N/A", safety_net_service)) %>%
  count(safety_net_service) %>%
  mutate(percentage = n / sum(n) * 100, legend_label = paste0(safety_net_service, " : ", n, " (", round(percentage, 1), "%)"))

ggplot(pie_data, aes(x = "", y = n, fill = legend_label)) + geom_bar(stat = "identity", width = 1) + coord_polar("y", start = 0) + theme_void() + scale_fill_brewer(palette = "Set2") + labs(title = "Comparison of Chinese Churches as Safety Nets", subtitle = "During the Japanese Occupation of Hong Kong", fill = "Safety Net Service Category") + theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 16), plot.subtitle = element_text(hjust = 0.5, size = 12, color = "gray40"), legend.position = "right", legend.title = element_text(face = "bold"), legend.text = element_text(size = 11))

```